Ab Initio Effective One-Electron Potentials Revisited: General Theory and Example Applications for Charge-Transfer Energy Calculations

### 2019-05-24 papers/2_oep_technique/paper-r0.tex
Group II of terms for CT energy described

\paragraph{Group (ii).}
\paragraph{Group (iii).}
\paragraph{Final OEP-based form of the coupling constant.}
Gathering the results from Eqs. the coupling constant
can be rewritten as follows
%
\begin{equation} \label{e:ct-murell-etal.vin.oep}
 V_{in}^{A\rightarrow B}
       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;.
\end{equation}
%
Note that the only approximations made on it so fare was the application of density fitting
and resolution of identity. If the associated basis sets are sufficiently large, the errors
due to this approximation can be minimal and negligible.

>>>>

\paragraph{Group (ii).}
The term belonging to this group can be considered as a sum of interaction
energies between the total charge density distribution of molecule $B$
and the partial density $\rho_{ik}({\bf r})$ of molecule $A$,
weighted by the overlap integrals $S_{nk}$. Using the distributed multipole 
expansion this can be written as
%
\begin{equation} \label{e:ct.group-2}
      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
 \cong - \sum_{k\in A}^{\rm Occ} S_{nk} \rho_{ki}^A \odot \rho^B_{\rm tot} \;.
\end{equation}
%
Consider now a situation in which the localized occupied MO's are used.
If the multipole series expansions as truncated
at the distributed charges $q_i$ and dipole moments ${\BM{\upmu}_i}$, that are
located at the centroids ${\bf r}_i = \tBraKet{i}{\hat{\bf r}}{i}$, 
then the interaction energy can be written approximately as
%
\begin{multline}
 \rho_{ki}^A \odot \rho^B_{\rm tot} 
 \approx 
 q_{ki} 
 \left[
  %\sum_{i\in A}^{\rm LC} 
 \sum_{y\in B}^{\rm AT}
  \frac{Z_y}{\vert {\bf r}_y - {\bf r}_i \vert } 
 +
 \sum_{j\in B}^{\rm LC}
  \frac{q_j}{\vert {\bf r}_j - {\bf r}_i \vert } 
 \right] \\
 + \text{ dipole terms } \;.
\end{multline}
%
Note however that $q_{ik} = -\delta_{ik}$ and $q_j = -1$. Moreover, 
if the localized orbitals are perfectly spherical distributed dipole moments vanish.
This means that Eq.~\eqref{e:ct.group-2} can be approximated as follows
%
\begin{equation}
      - \sum_{k\in A}^{\rm Occ} S_{nk} \tBraKet{k}{\hat{v}^B_{\rm tot} }{i} 
 \approx S_{ni}  \left[
  %\sum_{i\in A}^{\rm LC} 
 \sum_{y\in B}^{\rm AT}
  \frac{Z_y}{ r_{yi} } 
 -
 \sum_{j\in B}^{\rm LC}
  \frac{1}{r_{ji}} 
 \right] \;.
\end{equation}
%
Therefore, only overlap integrals are needed and relative distances between
atomic and LC's which leads to a great reduction of the evaluation cost
as compared to the original equation or the full multipole expansion.
We shall refer to this approximation as to
the spherical overlap approximation (SOA). Note that in order for this approximation to be valid,
occupied molecular orbitals need to be localized.

%If localized occupied MO's are used and only distributed charges ar, then $\rho_{ki} \approx \delta_{ki}$

\paragraph{Group (iii).}
\paragraph{Final OEP-based form of the coupling constant.}
Gathering the results from Eqs. the coupling constant
can be rewritten as follows
%
\begin{equation} \label{e:ct-murell-etal.vin.oep}
 V_{in}^{A\rightarrow B}
       = \sum_{\eta\in B}^{\rm RI} V^B_{n\eta} S_{\eta i} \;.
\end{equation}
%
Note that the only approximations made on it so fare was the application of density fitting
and resolution of identity. If the associated basis sets are sufficiently large, the errors
due to this approximation can be minimal and negligible.
